### 2021-11-21 thesis/tex/oscillations.tex
update

% \input{commands}
% \begin{document}

@@

We construct the PMNS matrix by parametrizing it as as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12} \nonumber \\
     &= \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
         U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
         U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
      \end{pmatrix} \nonumber \\
      & = \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix} \nonumber \\
 &=\begin{pmatrix}c_{12} c_{13} & s_{12} c_{13} & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 
        -s_{12} c_{23}-c_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{12} c_{23}-s_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & s_{23} c_{13} \\ 
        s_{12} s_{23}-c_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}}  & -c_{12} s_{23}-s_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{23} c_{13}
    \end{pmatrix}\,,
\end{align}
where $s_{ij} = \sin(\theta_{ij})$, $c_{ij} = \cos(\theta_{ij})$, $0 \leq \theta_{ij} \leq \pi/2$ and $ 0 \leq \delta_\mathrm{CP} \leq 2\pi$

By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure. We are now down to four degrees of freedom in the three dimensional case: three angles
%TODO: Look CKM wikipedia, good explanation. Mention that it is not guaranteed that the mixing matrix that we only have one phase. We could Have three due to Dirac/majorana. Majorana phases are not relevant for oscillations because oscillations are not violating lepton number, but include them in PMNS.
of the form $s_{ij}, c_{ij}$, and one phase of the form $e^{i\delta}$. The angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ mass eigenstate mixes with the $\nu_j$ mass eigenstate. Thus we refer to $\theta_{ij}$ as the mixing angles. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates $\nu_i$ that we allowed to mix. More on this in Section~\ref{sec:prop}.

@@

From the expression for the weak charged current in~\ref{eq:j_CC3}, we see that the summation over the mass index $k$ with the mixing matrix elements constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^\dagger$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^\dagger_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the vacuum Hamiltonian. 
The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse? We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^\dagger \ket{\nu_j} \nonumber \\
                              &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
                              &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
                 &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
                 &= \sum_{k,\b} U^\dagger_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) &= \sum_{k,j} {\bra{\nb}\ket{\na (t)}}^2 \nonumber \\
    &= \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\approx 1$, and $\abs{\vec{p}} \approx E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{{\vec{p}}^2 + m_k^2} \nonumber \\
        &= {\vec{p}}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
\end{align}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
\begin{align}
    %TODO: check abs
    P_{\ab}(t) = \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)t/2E}\,.
\end{align}
Now, our approximation $v\approx 1$ implies $x\approx t $, thus
\begin{align}
    P_{\ab}(x) &= \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E})\,,
\end{align}
where we in the last step have defined the \emph{mass-squared difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible to measure the absolute mass $m_k$ through neutrino oscillations. We have
to rely on other types of experiments to get us at least one absolute mass, from which we then can calculate the remaining.
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^\dagger = \delta_{\a \b}$ yields 
\begin{align}
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger]
\end{align}
%TODO: include dcp
We have
\begin{align}\label{eq:Pab} 
    %TODO: include dcp and don't write so many lines
    P_{\ab} &= \sum_{k,j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} +\sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} + 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}
which is the probability of neutrino vacuum oscillations.
The calculation for antineutrinos yield the same result if one continues to assume the realness of the mixing matrix.

Experimental measurements show that $\theta_{13}$ is the smallest of the mixing angles~\cite{nufit}, causing 
$\nm$ to primarily mix into $\nt$. The mixing matrix from~\ref{PMNS_def} takes the values
\begin{align}\label{eq:Uvalues}
   U = \begin{pmatrix}
       U_{e 1} & U_{e2} & U_{e3} \\
       U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
       U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
   \end{pmatrix} 
   = \begin{pmatrix}
       0.825 & 0.545 & 0.149 \\
       -0.455 & 0.485 & 0.746 \\
       0.334 & -0.684 & 0.649
   \end{pmatrix} \,,
\end{align}
which use to we plot the $P_{\mu \beta}$ probabilities in Fig.~\ref{fig:vac_osc}. Here, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12000}{\km}\footnote{This baseline
is approximately the Earth diameter, and will be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a suppressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
The smallness of $U_{e3}$ reduces the effect of $\dm[21]$ mass-squared difference,
allowing $\dm[31]$ to drive the oscillations. 
The most important mixing matrix 
elements for $P_{\a\b}$ are $U_{\a3}, U_{\a1}$ and $U_{\b3}, U_{\b1}$. %TODO: write whole 4 product here
%instead of U,U and U,U

@@

The three-neutrino oscillation Hamiltonian from Eq.~\eqref{eq:H_3gen} %TODO: derive this
now reduces to 
\begin{align}
    H_F &= \frac{1}{2E}(U M^2 U^\dagger + A) \nonumber \\
        &= \frac{1}{2E}\left[U \begin{pmatrix}
            0 & 0 & 0 \\
            0 & \dm_{21} & 0 \\
            0 & 0 & 0
        \end{pmatrix} U^\dagger\right] + \sqrt{2}G_F N_e \begin{pmatrix}
            1 & 0 & 0 \\
            0 & 0 & 0 \\
            0 & 0 & 0
        \end{pmatrix}\,. 
\end{align}
The rotation matrices $R_{ij}$ that constitute $U$ according to Eq.~\eqref{PMNS_def} can be simplified. The rotation $R_{13}$ will not affect our mass matrix, since the former acts in the $1-3$ plane, and the latter sits in the $1-2$ plane. Thus, the reduced mass matrix $\mathbb{M}_{12}$ commutes with $R_{13}$. Analogously, the  rotation matrix $R_{23}$ also commutes with $\mathbb{M}_{12}$.
Moreover, since $SO(3)$ is abelian, its generators commute, i.e.
\begin{align}
    \commutator{R_{ij}}{R_{kl}} = R_{ij}R_{kl} - R_{kl}R_{ij} = 0\,.
\end{align}
Thus,
\begin{align}
    U \mathbb{M}_{12} U^\dagger &=
    R_{23}R_{13}R_{12} \mathbb{M}_{12}  R_{12}^\dagger R_{13}^\dagger  R_{23}^\dagger \nonumber \\
    &= R_{12}R_{23}R_{13} \mathbb{M}_{12}  R_{13}^\dagger R_{23}^\dagger R_{12}^\dagger   \nonumber \\
    &= R_{12}\mathbb{M}_{12} R_{23} R_{13}   R_{13}^\dagger R_{23}^\dagger R_{12}^\dagger  \,.
\end{align}
And due to the fact that the rotation matrices are unitary,
\begin{align}
    R\,R^\dagger = \mathbb{I}\,,
\end{align}
we have that 
\begin{align}
    U \mathbb{M}_{12} U^\dagger &= R_{12}\mathbb{M}_{12}R_{23} R_{13}   R_{13}^\dagger R_{12}^\dagger  R_{23}^\dagger \nonumber \\
    &= R_{12}\mathbb{M}_{12}R_{23} R_{23}^\dagger  R_{12}^\dagger \nonumber \\
    &= R_{12}\mathbb{M}_{12}  R_{12}^\dagger\,.
\end{align}
We have now completely removed the dependence of the mixing angle between the $\nu_1$ and $\nu_3$ mass eigenstates, $\theta_{13}$, and the dependece of the mixing angle between the $\nu_2$ and $\nu_3$ mass eigenstates, $\theta_{23}$.
We can see this clearly by writing our reduced mixing matrix, which is only constituted by the rotation in the $1-2$ plane, $R_{12}$.
\begin{align}
    U &= R_{12}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1 \end{pmatrix} 
\end{align}

@@

While two-neutrino flavor oscillation can only be assumed to hold in regions where mixing into the third flavor can be considered negligible,
the fact that the mixing angle in matter depends on the squared-mass difference, and vice versa, gives us an interesting conclusion which is 
still valid for the full three neutrino oscillations. We previously saw that the oscillation amplitudes were entirely dependant on
the mixing angles, and the oscillation frequencies were entirely dependant on the squared-mass difference along with energy, and trajectory. After 
introducing the matter potentials, we now see that the matter effect mixes the mixing matrix elements with the masses and the energy. 
%TODO: maybe 2 vs 3 plot

>>>>

@@

We construct the PMNS matrix by parametrizing it as as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12}P_{Maj} \nonumber \\
     &= \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
         U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
         U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
      \end{pmatrix} \nonumber \\
      & = \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix}
\begin{pmatrix}1 & 0 & 0 \\ 0 & e^{i \lambda{2}} & 0 \\ 0 & 0 & e^{i \lambda{3}}\end{pmatrix}
\nonumber \\
 &=\begin{pmatrix}c_{12} c_{13} & s_{12} c_{13} & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 
        -s_{12} c_{23}-c_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{12} c_{23}-s_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & s_{23} c_{13} \\ 
        s_{12} s_{23}-c_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}}  & -c_{12} s_{23}-s_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{23} c_{13}
    \end{pmatrix}
    \begin{pmatrix}1 & 0 & 0 \\ 0 & e^{i \lambda{2}} & 0 \\ 0 & 0 & e^{i \lambda{3}}\end{pmatrix}\,,
\end{align}
where $s_{ij} = \sin(\theta_{ij})$, $c_{ij} = \cos(\theta_{ij})$, $0 \leq \theta_{ij} \leq \pi/2$ and $ 0 \leq \lambda{i},\delta_\mathrm{CP} \leq 2\pi$.

By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure: one term in the form of a Dirac CP-violating phase $\delta_\mathrm{CP}$.
However, if the neutrinos are Majorana fermions, the rephasing of the left-handed fields would cause the neutrino masses to become complex.
Thus, we include two Majorana phases $\lambda{2},\lambda_3$ in the PMNS matrix. These two phases are detectable in processes that violate lepton number, such as neutrinoless double $\beta$ decay.
Majorana phases are not detectable in neutrino oscillation experiments because oscillations are not violating lepton number.

We are now down to six degrees of freedom in the three dimensional case: three angles
of the form $s_{ij}, c_{ij}$, and three phase of the form $e^{i\delta_\mathrm{CP}}$ and $e^{\lambda{i}}$. The angle $\theta_{ij}$ specifies the amount of rotation in the $i-j$ plane, allowing an $i$ eigenstate to mix with a $j$ eigenstate. The physical interpretation of
a mixing angle is then the degree to which the $\nu_i$ mass eigenstate mixes with the $\nu_j$ mass eigenstate. Thus we refer to $\theta_{ij}$ as the mixing angles. A non-zero mixing angle will then 
cause the flavor eigenstate $\nu_\alpha$ to be a superposition of the mass eigenstates $\nu_i$ that we allowed to mix. More on this in Section~\ref{sec:prop}.

@@

From the expression for the weak charged current in~\ref{eq:j_CC3}, we see that the summation over the mass index $k$ with the mixing matrix elements constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^\dagger$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^\dagger_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the vacuum Hamiltonian. 
The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse? We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^\dagger \ket{\nu_j} \nonumber \\
                              &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
                              &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
                 &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
                 &= \sum_{k,\b} U^\dagger_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) &= \sum_{k,j} \abs{{\bra{\nb}\ket{\na (t)}}}^2 \nonumber \\
    &= U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\approx 1$, and $\abs{\vec{p}} \approx E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{{\vec{p}}^2 + m_k^2} \nonumber \\
        &= {\vec{p}}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
\end{align}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
\begin{align}
    P_{\ab}(t) = \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)t/2E}\,.
\end{align}
Now, our approximation $v\approx 1$ implies $x\approx t $, thus
\begin{align}
    P_{\ab}(x) &= \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} \abs{U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E})\,,
\end{align}
where we in the last step have defined the \emph{mass-squared difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible to measure the absolute mass $m_k$ through neutrino oscillations. We have
to rely on other types of experiments to get us at least one absolute mass, from which we then can calculate the remaining.
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^\dagger = \delta_{\a \b}$ yields 
\begin{align}
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger]
\end{align}
We have
\begin{align}\label{eq:Pab} 
    P_{\ab} &= \sum_{k,j} \abs{U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \sum_{k \neq j} \abs{U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger] +\sum_{k \neq j} \abs{U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            %&= \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger] + 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            %&= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 4\sum_{k > j} \Re[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)} \nonumber \\
            &\hspace{1.2cm}+ 2\sum_{k > j} \Im[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}
which is the probability of neutrino vacuum oscillations.
A similar calculation for antineutrinos yield  
\begin{align}\label{eq:Pab} 
    P_{\ab} 
            &= \delta_{\a \b} - 4\sum_{k > j} \Re[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)} \nonumber \\
            &\hspace{1.2cm}- 2\sum_{k > j} \Im[U_{\a k} ^ \dagger U_{\b k}U_{\b j}^\dagger U_{\a j}] \sin{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}

Experimental measurements show that $\theta_{13}$ is the smallest of the mixing angles~\cite{nufit}, causing 
$\nm$ to primarily mix into $\nt$. Using values from~\cite{nufit}, but taking $\delta_\mathrm{CP} = 0$ for simplicity,
the mixing matrix from~\ref{PMNS_def} takes the values
\begin{align}\label{eq:Uvalues}
   U = \begin{pmatrix}
       U_{e 1} & U_{e2} & U_{e3} \\
       U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
       U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
   \end{pmatrix} 
   = \begin{pmatrix}
       0.825 & 0.545 & 0.149 \\
       -0.455 & 0.485 & 0.746 \\
       0.334 & -0.684 & 0.649
   \end{pmatrix} \,,
\end{align}
which use to we plot the $P_{\mu \beta}$ probabilities in Fig.~\ref{fig:vac_osc}. Here, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12000}{\km}\footnote{This baseline
is approximately the Earth diameter, and will be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a suppressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
The smallness of $U_{e3}$ reduces the effect of $\dm[21]$ mass-squared difference,
allowing $\dm[31]$ to drive the oscillations. 
The most important mixing matrix 
elements for $P_{\a\b}$ are $U_{\a3}, U_{\a1}$ and $U_{\b3}, U_{\b1}$. %TODO: write whole 4 product here
%instead of U,U and U,U

@@

The three-neutrino oscillation Hamiltonian from Eq.~\eqref{eq:H_3gen}
now reduces to 
\begin{align}
    H_F &= \frac{1}{2E}(U M^2 U^\dagger + A) \nonumber \\
        &= \frac{1}{2E}\left[U \begin{pmatrix}
            0 & 0 & 0 \\
            0 & \dm_{21} & 0 \\
            0 & 0 & 0
        \end{pmatrix} U^\dagger\right] + \sqrt{2}G_F N_e \begin{pmatrix}
            1 & 0 & 0 \\
            0 & 0 & 0 \\
            0 & 0 & 0
        \end{pmatrix}\,. 
\end{align}
The rotation matrices $R_{ij}$ that constitute $U$ according to Eq.~\eqref{PMNS_def} can be simplified. The rotation $R_{13}$ will not affect our mass matrix, since the former acts in the $1-3$ plane, and the latter sits in the $1-2$ plane. Thus, the reduced mass matrix $\mathbb{M}_{12}$ commutes with $R_{13}$. Analogously, the  rotation matrix $R_{23}$ also commutes with $\mathbb{M}_{12}$.
Moreover, since $SO(3)$ is abelian, its generators commute, i.e.
\begin{align}
    \commutator{R_{ij}}{R_{kl}} = R_{ij}R_{kl} - R_{kl}R_{ij} = 0\,.
\end{align}
Thus,
\begin{align}
    U \mathbb{M}_{12} U^\dagger &=
    R_{23}R_{13}R_{12} \mathbb{M}_{12}  R_{12}^\dagger R_{13}^\dagger  R_{23}^\dagger \nonumber \\
    &= R_{12}R_{23}R_{13} \mathbb{M}_{12}  R_{13}^\dagger R_{23}^\dagger R_{12}^\dagger   \nonumber \\
    &= R_{12}\mathbb{M}_{12} R_{23} R_{13}   R_{13}^\dagger R_{23}^\dagger R_{12}^\dagger  \,.
\end{align}
And due to the fact that the rotation matrices are unitary,
\begin{align}
    R\,R^\dagger = \mathbb{I}\,,
\end{align}
we have that 
\begin{align}
    U \mathbb{M}_{12} U^\dagger &= R_{12}\mathbb{M}_{12}R_{23} R_{13}   R_{13}^\dagger R_{12}^\dagger  R_{23}^\dagger \nonumber \\
    &= R_{12}\mathbb{M}_{12}R_{23} R_{23}^\dagger  R_{12}^\dagger \nonumber \\
    &= R_{12}\mathbb{M}_{12}  R_{12}^\dagger\,.
\end{align}
We have now completely removed the dependence of the mixing angle between the $\nu_1$ and $\nu_3$ mass eigenstates, $\theta_{13}$, and the dependece of the mixing angle between the $\nu_2$ and $\nu_3$ mass eigenstates, $\theta_{23}$.
We can see this clearly by writing our reduced mixing matrix, which is only constituted by the rotation in the $1-2$ plane, $R_{12}$.
\begin{align}
    U &= R_{12}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1 \end{pmatrix} 
\end{align}

@@

While two-neutrino flavor oscillation can only be assumed to hold in regions where mixing into the third flavor can be considered negligible,
the fact that the mixing angle in matter depends on the squared-mass difference, and vice versa, gives us an interesting conclusion which is 
still valid for the full three neutrino oscillations. We previously saw that the oscillation amplitudes were entirely dependant on
the mixing angles, and the oscillation frequencies were entirely dependant on the squared-mass difference along with energy, and trajectory. After 
introducing the matter potentials, we now see that the matter effect mixes the mixing matrix elements with the masses and the energy.
